Cosmology with photometric surveys: constraints with redshift tomography of non--Gaussian statistics

### 2016-03-28 photoz_paper/paper.tex
create new figures

\subsection{Cosmic shear simulations}
\label{sec:shearsim}
We review the procedure used for generating mock shear catalogs. We consider a fiducial flat $\Lambda$CDM universe with $(h,\Omega_m,\Omega_\Lambda,\Omega_b,w,\sigma_8,n_s)=(0.72,0.26,0.74,0.046,-1,0.8,0.96)$. We examine different variations of the $\bb{p}=(\Omega_m,w,\sigma_8)$ triplet and run one $N$--body simulation for each choice of $\bb{p}$, using the public code \ttt{Gadget2} \citep{Gadget2}. The simulations have a comoving box size of $L_b=260\,{\rm Mpc}/h$ and contain $512^3$ dark matter particles, which correspond to a mass resolution of $M_p\approx 10^{10}M_{\rm sun}$ per particle. The three dimensional outputs of the $N$--body simulations are sliced in sequences of two dimensional lenses $120 \,{\rm Mpc}$ thick, which are lined up perpendicular to the line of sight between the observer on Earth and a source at redshift $z_s$. We make use of the multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} to trace the deflections of light rays originating at $z=0$ through the system of lenses until $z$. To accomplish this task, we make use of the \ttt{LensTools} \citep{LensTools-ASCL,LensTools-paper} implementation of the multi--lens--plane algorithm. An observed galaxy position $\pmb{\theta}$ on the sky today corresponds to a real galaxy angular position $\pmb{\beta}(\pmb{\theta},z_s)$, which can be calculated using the \ttt{LensTools} pipeline by solving the lensing ODE up to redshift $z_s$. The Jacobian of $\pmb{\beta}(\pmb{\theta},z_s)$ is a $2\times 2$ matrix that contains information about the cosmic shear field at $\pmb{\theta}$ integrated along the line of sight. 
%
\begin{equation}
\label{meth:sheareqn}
\frac{\partial\beta_i(\pmb{\theta},z_s)}{\partial \theta_j} = 
\begin{pmatrix}
1-\kappa(\pmb{\theta})-\gamma^1(\pmb{\theta}) & -\gamma^2(\pmb{\theta}) \\
-\gamma^2(\pmb{\theta}) & 1-\kappa(\pmb{\theta})+\gamma^1(\pmb{\theta})\\
\end{pmatrix}
\end{equation}  
%
The quantities that appear in equation (\ref{meth:sheareqn}) are the convergence $\kappa$, which is the source magnification due to lensing, and the cosmic shear $\pmb{\gamma}$, which is a measurement of the source ellipticity due to lensing from Large Scale Structure, assuming the non lensed shape is a circle. 
We simulate $N_g = 10^6$ random galaxy positions $\{\pmb{\theta}_g\}$ distributed uniformly in a field of view of size $\theta_{\rm FOV}^2=(3.5{\rm deg})^2$, which correspond to a galaxy angular density of $n_g=22\,\rm{arcmin}^{-2}$. The galaxy have a distribution in redshift which mimics the one expected in the LSST survey
\begin{equation}
n(z) = n_0\left(\frac{z}{z_0}\right)^2\exp\left(-\frac{z}{z_0}\right)
\end{equation}  
%
with $z_0=0.3$ and $n_0$ a normalization constant fixed so that $n(z)$ integrates to the total number of galaxies $N_g$. For each galaxy we compute the cosmic shear at $\pmb{\theta}_g$ using equation (\ref{meth:sheareqn}), producing a shear catalog $\{\pmb{\gamma}_g\}$. Different random realizations of a shear catalog $\{\pmb{\gamma}_g\}_r$ can be obtained rotating and periodically shifting the Large Scale Structure in the $N$--body snapshots according to the procedure explained in \citep{PetriVariance}. We produce $N_r=16000$ pseudo--independent realizations of the shear catalog $\{\pmb{\gamma}_g\}$. These shear realization cover the total survey area of LSST 10 times.
We repeat the above procedure for $P=100$ different combinations of the parameter triplet $\bb{p}$, sampled according to a Latin hypercube scheme. The sampling procedure is the same described in \citep{CFHTMink,CFHTPeaks}.

%ADD Simulation specifications

For each of these simulations, the $N$--body initial conditions are generated using the same random seed. In addition to this simulations, we produce simulated shear catalogs for a fiducial $\Lambda$CDM universe with $\bb{p}_0=(0.26,-1,0.8)$. In this case the randomization procedure is based on 5 independent $N$--body simulations, and the same number of pseudo--independent catalog realizations is produced.

@@

In addition to shape noise contributions to the observed galaxy ellipticity we consider photometric redshift errors as an additional contamination source in the simulated catalogs. In photometric surveys such as LSST, the source redshift $z_s$ is estimated measuring the source luminosity in a finite small set of optical frequency bands. Using this compressed luminosity information rather than the full spectrum introduces biases in redshift estimation. Forward modeling of the cosmic shear using the procedure described in \S~\ref{sec:shearsim}, as well as the shape noise contributions, assume a correct redshift distributions $n(z)$. An incorrect binning of observed galaxy redshifts according to the measured photometric distribution $n_p(z_p)$ can propagate the redshift measurement errors all the way to cosmological parameter constraints if the latter take advantage of redshift tomography. One of the goals of this work is to evaluate the size of this effect, assuming photometric redshift errors (photo-$z$) are left uncorrected. We model effect of photo-$z$ errors as a constant bias term $b_{\rm ph}(z_s)$ plus a random Gaussian component with variance $\sigma_{\rm ph}(z_s)$
\begin{equation}
\label{meth:photoz-correction}
z_p(z_s) = z_s + b_{\rm ph}(z_s) + \sigma_{\rm ph}(z_s)\mathcal{N}(0,1)   
\end{equation}
%
where $\mathcal{N}(0,1)$ is the standard normal distribution. We bin the $N_g$ galaxies in our simulated catalogs in 5 redshift bins $\bar{z}_b$, $b=1...5$. Several models have been proposed in the literature for the photometric bias $b_{\rm ph}(z_s)$ (see for example \citep{Huterer2006}) and variance $\sigma_{\rm ph}(z_s)$ (see for example \citep{LSSTSciBook}). We make use of the photo-$z$ bias and variance values contained in the LSST Science Book \citep{LSSTSciBook}. The bin sizes are adjusted so that each of them contains the same number of galaxies. The galaxy distribution we assumed, along with the binning choice is shown in Figure \ref{fig:galdistr}

@@

@@

>>>>

\subsection{Cosmic shear simulations}
\label{sec:shearsim}
We review the procedure used for generating mock shear catalogs. We consider a fiducial flat $\Lambda$CDM universe with $(h,\Omega_m,\Omega_\Lambda,\Omega_b,w,\sigma_8,n_s)=(0.72,0.26,0.74,0.046,-1,0.8,0.96)$. We examine different variations of the $\bb{p}=(\Omega_m,w,\sigma_8)$ triplet and run one $N$--body simulation for each choice of $\bb{p}$, using the public code \ttt{Gadget2} \citep{Gadget2}. The simulations have a comoving box size of $L_b=260\,{\rm Mpc}/h$ and contain $512^3$ dark matter particles, which correspond to a mass resolution of $M_p\approx 10^{10}M_{\rm sun}$ per particle. The three dimensional outputs of the $N$--body simulations are sliced in sequences of two dimensional lenses $120 \,{\rm Mpc}$ thick, which are lined up perpendicular to the line of sight between the observer on Earth and a source at redshift $z_s$. We make use of the multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} to trace the deflections of light rays originating at $z=0$ through the system of lenses until $z$. To accomplish this task, we make use of the \ttt{LensTools} \citep{LensTools-ASCL,LensTools-paper} implementation of the multi--lens--plane algorithm. An observed galaxy position $\pmb{\theta}$ on the sky today corresponds to a real galaxy angular position $\pmb{\beta}(\pmb{\theta},z_s)$, which can be calculated using the \ttt{LensTools} pipeline by solving the lensing ODE up to redshift $z_s$. The Jacobian of $\pmb{\beta}(\pmb{\theta},z_s)$ is a $2\times 2$ matrix that contains information about the cosmic shear field at $\pmb{\theta}$ integrated along the line of sight. 
%
\begin{equation}
\label{meth:sheareqn}
\frac{\partial\beta_i(\pmb{\theta},z_s)}{\partial \theta_j} = 
\begin{pmatrix}
1-\kappa(\pmb{\theta})-\gamma^1(\pmb{\theta}) & -\gamma^2(\pmb{\theta}) \\
-\gamma^2(\pmb{\theta}) & 1-\kappa(\pmb{\theta})+\gamma^1(\pmb{\theta})\\
\end{pmatrix}
\end{equation}  
%
The quantities that appear in equation (\ref{meth:sheareqn}) are the convergence $\kappa$, which is the source magnification due to lensing, and the cosmic shear $\pmb{\gamma}$, which is a measurement of the source ellipticity due to lensing from Large Scale Structure, assuming the non lensed shape is a circle. 
We simulate $N_g = 10^6$ random galaxy positions $\{\pmb{\theta}_g\}$ distributed uniformly in a field of view of size $\theta_{\rm FOV}^2=(3.5{\rm deg})^2$, which correspond to a galaxy angular density of $n_g=22\,\rm{arcmin}^{-2}$. The galaxy have a distribution in redshift which mimics the one expected in the LSST survey
\begin{equation}
n(z) = n_0\left(\frac{z}{z_0}\right)^2\exp\left(-\frac{z}{z_0}\right)
\end{equation}  
%
with $z_0=0.3$ and $n_0$ a normalization constant fixed so that $n(z)$ integrates to the total number of galaxies $N_g$. For each galaxy we compute the cosmic shear at $\pmb{\theta}_g$ using equation (\ref{meth:sheareqn}), producing a shear catalog $\{\pmb{\gamma}_g\}$. Different random realizations of a shear catalog $\{\pmb{\gamma}_g\}_r$ can be obtained rotating and periodically shifting the Large Scale Structure in the $N$--body snapshots according to the procedure explained in \citep{PetriVariance}. We produce $N_r=16000$ pseudo--independent realizations of the shear catalog $\{\pmb{\gamma}_g\}$. These shear realization cover the total survey area of LSST 10 times.
We repeat the above procedure for $P=100$ different combinations of the parameter triplet $\bb{p}$, sampled according to a Latin hypercube scheme. The sampling procedure is the same described in \citep{CFHTMink,CFHTPeaks}. The parameter space sampling we adopted for our simulations is shown in Figure \ref{fig:simdesign}

\begin{figure*}
\includegraphics[scale=0.4]{Figures/design.eps}
\caption{Parameter space sampling}
\label{fig:simdesign}
\end{figure*}
%
For each of the parameter choices in Figure \ref{fig:simdesign}, the $N$--body initial conditions are generated using the same random seed. In addition to this simulations, we produce simulated shear catalogs for a fiducial $\Lambda$CDM universe with $\bb{p}_0=(0.26,-1,0.8)$. In this case the randomization procedure is based on 5 independent $N$--body simulations, and the same number of pseudo--independent catalog realizations is produced.

@@

In addition to shape noise contributions to the observed galaxy ellipticity we consider photometric redshift errors as an additional contamination source in the simulated catalogs. In photometric surveys such as LSST, the source redshift $z_s$ is estimated measuring the source luminosity in a finite small set of optical frequency bands. Using this compressed luminosity information rather than the full spectrum introduces biases in redshift estimation. Forward modeling of the cosmic shear using the procedure described in \S~\ref{sec:shearsim}, as well as the shape noise contributions, assume a correct redshift distributions $n(z)$. An incorrect binning of observed galaxy redshifts according to the measured photometric distribution $n_p(z_p)$ can propagate the redshift measurement errors all the way to cosmological parameter constraints if the latter take advantage of redshift tomography. One of the goals of this work is to evaluate the size of this effect, assuming photometric redshift errors (photo-$z$) are left uncorrected. We model effect of photo-$z$ errors as a constant bias term $b_{\rm ph}(z_s)$ plus a random Gaussian component with variance $\sigma_{\rm ph}(z_s)$
\begin{equation}
\label{meth:photoz-correction}
z_p(z_s) = z_s + b_{\rm ph}(z_s) + \sigma_{\rm ph}(z_s)\mathcal{N}(0,1)   
\end{equation}
%
where $\mathcal{N}(0,1)$ is the standard normal distribution. We bin the $N_g$ galaxies in our simulated catalogs in 5 redshift bins $\bar{z}_b$, $b=1...5$. Several models have been proposed in the literature for the photometric bias $b_{\rm ph}(z_s)$ (see for example \citep{Huterer2006}) and variance $\sigma_{\rm ph}(z_s)$ (see for example \citep{LSSTSciBook}). We make use of the photo-$z$ bias and variance values contained in the LSST Science Book \citep{LSSTSciBook}, which we show in Figure \ref{fig:lsstscibook}. The bin sizes are adjusted so that each of them contains the same number of galaxies. The galaxy distribution we assumed, along with the binning choice is shown in Figure \ref{fig:galdistr}

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/scibook.eps}
\caption{LSST Science Book}
\label{fig:lsstscibook}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/constraints.eps}
\caption{Constraints in $(\Omega_m,w)$ space}
\label{fig:constraintsOm-w}
\end{figure}
